---
title: "Interactive charts in Quarto revealjs"
subtitle: "Four ways to build the same interactive slider chart"
author: "Demo deck"
date: 2026-08-04
format:
  revealjs:
    theme: simple
    embed-resources: true
    slide-number: c
    # A bigger slide canvas is the YAML-native way to shrink type: Reveal lays
    # out in these coordinates and scales the whole slide to fit the screen, so
    # headings and body text come down together. 1500x940 is ~77% of 1150x720.
    width: 1500
    height: 940
---

```{r}
#| label: setup
#| include: false

library(tidyverse)

dir.create("figs", showWarnings = FALSE)
set.seed(2026)

# One palette, shared by the R charts and by the two JavaScript charts,
# which hard-code the same hex values.
pal <- list(accent = "#c2410c", spell = "#0f766e", gold = "#ca8a04",
            text = "#1f2328", subtle = "#6b7280", border = "#d8dce1")

# Resident income tax schedule from 1 July 2024, in dollars. The charts apply
# this rather than a single flat rate, so the property price now matters:
# 100 index points = PRICE dollars.
PRICE <- 100000
brackets <- tibble(lo   = c(0, 18200, 45000, 135000, 190000),
                   rate = c(0, 0.16,  0.30,  0.37,   0.45))

schedule_tax <- function(income) {
  hi <- c(brackets$lo[-1], Inf)
  tax <- sum(pmax(0, pmin(income, hi) - brackets$lo) * brackets$rate)
  tax + min(0.02 * income, 0.10 * max(0, income - 24276))
}

dollars <- function(index) index / 100 * PRICE

# The rate slider reads as a marginal rate; it sets pre-gain taxable income to
# the bottom of the bracket where that rate applies.
income_at_rate <- function(tau) {
  i <- detect_index(brackets$rate, \(r) r >= tau - 1e-9)
  if (i == 0) last(brackets$lo) else brackets$lo[i]
}
```

## 1. Hand-written SVG and plain JavaScript

```{=html}
<style>
/* The only CSS in the deck, and both rules are mechanics rather than styling. */

/* Precomputed frames: eleven images stacked, one visible at a time. */
.frame-stack img { display: none; width: 100%; }
.frame-stack img.frame-on { display: block; }

/* The two hand-drawn SVG charts fill their column without overflowing the
   slide: both are a 640x610 viewBox, taller than a slide at full width. */
.flip-chart svg { width: 100%; height: auto; max-height: 575px; }
</style>
```

{{< include _flip-javascript.qmd >}}

::: aside
About 270 lines in `_flip-javascript.qmd`, of which roughly 15 are the
economics; the rest builds and moves SVG nodes. No libraries. Every slider tick
re-solves the model and mutates attributes on nodes that already exist, which is
why it is instant. The fading trail and the ghost line depend on which slider
moved last — trivial to track here, awkward everywhere else.
:::

## 2. Observable JS and Observable Plot

```{ojs}
//| output: false
// Quarto ships the Observable runtime, Plot, and Inputs. Nothing is fetched.
ink = ({accent: "#c2410c", spell: "#0f766e", gold: "#ca8a04",
        text: "#1f2328", subtle: "#6b7280", bg: "#ffffff"})

mortgageCost = function (rate, io) {
  let pmt = rate / (1 - Math.pow(1 + rate, -30));
  let bal = 1, sumIO = 0, sumPI = 0;
  for (let t = 1; t <= 10; t++) {
    const w = Math.pow(1.025, 10 - t);
    sumIO += rate * w;
    sumPI += rate * bal * w;
    bal -= pmt - rate * bal;
  }
  return io ? sumIO : sumPI;
}

// Personal income tax: the resident schedule from 1 July 2024, not a flat
// rate. P0 = 100 index points = PRICE dollars, so scale now matters.
PRICE = 100000
BR_LO = [0, 18200, 45000, 135000, 190000]
BR_RATE = [0, 0.16, 0.30, 0.37, 0.45]
scheduleTax = function (y) {
  let t = 0;
  for (let i = 0; i < BR_LO.length; i++) {
    const hi = (i + 1 < BR_LO.length) ? BR_LO[i + 1] : Infinity;
    t += Math.max(0, Math.min(y, hi) - BR_LO[i]) * BR_RATE[i];
  }
  return t + Math.min(0.02 * y, 0.10 * Math.max(0, y - 24276));
}
// The rate slider still reads as a marginal rate; it sets the investor's
// pre-gain taxable income to the bottom of the bracket where that rate
// applies, and the gain then stacks up through the schedule from there.
incomeAtRate = function (tau) {
  for (let i = 0; i < BR_LO.length; i++) if (BR_RATE[i] >= tau - 1e-9) return BR_LO[i];
  return BR_LO[BR_LO.length - 1];
}
dollars = (idx) => idx / 100 * PRICE

// The whole model, with nothing about pixels in it. Returns a function
// giving after-tax return for any before-tax return, because the schedule
// makes that relation piecewise linear rather than a single slope.
flip = function ({gain, lvr, rate, rent, cost, tau, delta, io, sd}) {
  const T = 10, P0 = 100, INFL = 0.025;
  const dutyNom = sd ? 0.05 * P0 : 0;
  const dutyReal = dutyNom * Math.pow(1 + INFL, T);
  const rentTot = rent * P0 * T;
  const cgNet = (1 - 0.02 * cost) * P0 * (1 + gain) - P0;
  const nir = rentTot -
    (0.055 * rentTot + 3 * (rent * P0 / 52) + 0.015 * P0 * T) * cost;
  const rentalTotal = nir - mortgageCost(rate, io) * lvr * P0;
  const E = cgNet + rentalTotal - dutyReal;

  // Rental result is deducted (or taxed) in each of the ten years; the
  // discounted gain lands once, in the year of sale.
  const other = incomeAtRate(tau);
  const rentalYr = dollars(rentalTotal / T);
  const taxBase = scheduleTax(other);
  const taxRent = scheduleTax(other + rentalYr);
  const taxAt = (e) => {
    const cg = e - rentalTotal + dutyReal;
    const gainTaxable = dollars((1 - delta) * (cg - dutyNom));
    return (T * (taxRent - taxBase) +
            scheduleTax(other + rentalYr + gainTaxable) - taxRent) / PRICE * 100;
  };
  const yAt = (eAnnual) => (eAnnual * T - taxAt(eAnnual * T)) / T;
  return {eYr: E / T, aYr: (E - taxAt(E)) / T, yAt: yAt, floorYr: flipFloor(yAt)};
}

// The curve is kinked, so the flip floor is bisected rather than solved.
flipFloor = function (yAt) {
  if (yAt(0) <= 0) return null;
  if (yAt(-4) > 0) return -4;
  let lo = -4, hi = 0;
  for (let k = 0; k < 44; k++) {
    const mid = (lo + hi) / 2;
    if (yAt(mid) > 0) hi = mid; else lo = mid;
  }
  return hi;
}
```

```{ojs}
//| output: false
clampDom = (v) => Math.min(Math.max(v, -4), 4)

// Same 640x610 frame, margins, ticks and annotations as version 1, so the two
// charts are pixel-identical. Note what that costs: more than half of this
// spec exists to switch Plot's defaults back off.
flipPlot = function (out) {
  const labelY = Math.min(Math.max(out.yAt(-3.15), -3.5), 3.5);
  const dx = clampDom(out.eYr), dy = clampDom(out.aYr);
  return Plot.plot({
    width: 640, height: 610,
    marginLeft: 50, marginRight: 14, marginTop: 14, marginBottom: 48,
    style: {color: ink.text, background: "transparent", fontSize: "13px"},
    x: {domain: [-4, 4], ticks: [-4, -3, -2, -1, 0, 1, 2, 3, 4],
        tickFormat: (d) => d + "%", tickSize: 0, tickPadding: 10, label: null},
    y: {domain: [-4, 4], ticks: [-4, -3, -2, -1, 1, 2, 3, 4],
        tickFormat: (d) => String(d), tickSize: 0, tickPadding: 6, label: null},
    marks: [
      out.floorYr === null ? null
        : Plot.rect([{x1: Math.max(out.floorYr, -4), x2: 0, y1: -4, y2: 4}],
                    {x1: "x1", x2: "x2", y1: "y1", y2: "y2",
                     fill: ink.gold, fillOpacity: 0.18}),
      Plot.ruleY([0], {stroke: ink.text, strokeOpacity: 0.35}),
      Plot.ruleX([0], {stroke: ink.text, strokeOpacity: 0.35}),
      Plot.line([[-4, -4], [4, 4]], {stroke: ink.subtle, strokeDasharray: "5 4"}),
      Plot.text([{x: 3.3, y: 2.8, t: "No tax"}],
                {x: "x", y: "y", text: "t", dy: -4, fill: ink.subtle, fontSize: 14}),
      out.floorYr === null ? null
        : Plot.text([{x: 0, y: -4, t: "flip\ninterval"}],
                    {x: "x", y: "y", text: "t", dx: -6, dy: -12,
                     textAnchor: "end", lineAnchor: "bottom",
                     lineHeight: 1.15, fill: ink.gold, fontSize: 14}),
      Plot.line(d3.range(0, 97).map(i => { const e = -4 + 8 * i / 96;
                                            return [e, out.yAt(e)]; }),
                {stroke: ink.accent, strokeWidth: 3, clip: true}),
      Plot.text([{x: -3.15, y: labelY, t: "After tax"}],
                {x: "x", y: "y", text: "t", dy: -17,
                 fill: ink.accent, fontSize: 14}),
      Plot.dot([[dx, dy]], {r: 7, fill: ink.spell, stroke: ink.bg,
                            strokeWidth: 2, clip: true}),
      Plot.text([{x: dx, y: dy, t: "Property\nspell"}],
                {x: "x", y: "y", text: "t", dx: -12, dy: -12,
                 textAnchor: "end", lineAnchor: "bottom",
                 lineHeight: 1.15, fill: ink.spell, fontSize: 14}),
      Plot.text(["Before-tax return, % a year"],
                {frameAnchor: "bottom", dy: 40, fill: ink.text, fontSize: 15}),
      Plot.text(["After-tax return, % a year"],
                {frameAnchor: "top-left", dx: 8, dy: 7, textAnchor: "start",
                 fill: ink.text, fontSize: 15})
    ]
  })
}
```

:::: {.columns}
::: {.column width="38%"}
::: {style="font-size:0.62em"}
```{ojs}
//| echo: false
viewof o_cg   = Inputs.range([10, 100], {value: 40, step: 5,   label: "Capital gain, %"})
viewof o_lvr  = Inputs.range([0, 100],  {value: 100, step: 5,  label: "LVR, %"})
viewof o_rate = Inputs.range([4, 9],    {value: 6, step: 0.25, label: "Mortgage rate, %"})
viewof o_rent = Inputs.range([2, 6],    {value: 4, step: 0.5,  label: "Rent, % of value"})
viewof o_tau  = Inputs.select([0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.47],
                              {value: 0.45, label: "Marginal tax rate",
                               format: (x) => (100 * x).toFixed(0) + "%"})
viewof o_delta = Inputs.range([0, 0.5], {value: 0.5, step: 0.05, label: "CGT discount"})
viewof o_io   = Inputs.toggle({value: true, label: "Interest only"})
```

```{ojs}
//| output: false
o_flipText = o_out.floorYr === null ? "No wedge, so no loss can flip." : "Losses up to **" + Math.abs(o_out.floorYr).toFixed(1) + "%** a year flip into after-tax profits."
o_spellText = " The spell returns **" + o_out.eYr.toFixed(1) + "%** before tax and **" + o_out.aYr.toFixed(1) + "%** after."
```

```{ojs}
//| echo: false
md`${o_flipText}${o_spellText}`
```
:::
:::

::: {.column width="60%" .flip-chart}
```{ojs}
//| echo: false
o_out = flip({gain: o_cg / 100, lvr: o_lvr / 100, rate: o_rate / 100,
              rent: o_rent / 100, cost: 1, tau: o_tau, delta: o_delta,
              io: o_io, sd: false})

flipPlot(o_out)
```
:::
::::

::: aside
Pixel-for-pixel the same chart as version 1, in about a third of the code: 30
lines of model with no pixels in it, plus a 55-line Plot spec. Left alone, Plot
draws a good chart that is not your chart, so much of that spec is
`tickSize: 0`, explicit tick lists, `label: null`. The trail and ghost line are
gone — "what changed since last time" is what a reactive dataflow discards.
:::

## 3. R computes, JavaScript only draws

```{r}
#| label: flip-grid
#| include: false

# The model, written where it can be read, reviewed and tested.
flip_pre_tax <- function(gain, lvr, rate, interest_only,
                         rent = 0.04, cost = 1, hold = 10, infl = 0.025,
                         p0 = 100) {
  # Each year's interest, carried to end-of-period real terms.
  weights <- (1 + infl)^(hold - seq_len(hold))
  payment <- rate / (1 - (1 + rate)^-30)

  balance <- if (interest_only) {
    rep(1, hold)
  } else {
    accumulate(seq_len(hold - 1), \(bal, i) bal - (payment - rate * bal), .init = 1)
  }

  rent_total <- rent * p0 * hold
  costs      <- (0.055 * rent_total + 3 * (rent * p0 / 52) + 0.015 * p0 * hold) * cost
  cg_net     <- (1 - 0.02 * cost) * p0 * (1 + gain) - p0

  # Before-tax payoff and the taxable gain. Tax enters later, and linearly.
  tibble(
    e_pre  = cg_net + rent_total - costs - sum(rate * balance * weights) * lvr * p0,
    cg_net = cg_net
  )
}

flip_grid <- expand_grid(
  gain = seq(0.10, 1.00, by = 0.10),
  lvr  = seq(0, 1, by = 0.10),
  rate = seq(0.04, 0.09, by = 0.005),
  io   = c(TRUE, FALSE)
) |>
  mutate(payoff = pmap(list(gain, lvr, rate, io), flip_pre_tax)) |>
  unnest(payoff) |>
  mutate(across(c(e_pre, cg_net), \(x) round(x, 4)))

ojs_define(flipgrid = flip_grid)
```

```{ojs}
//| output: false
// Two lines of arithmetic left in the browser, because after-tax return is
// linear in the tax parameters once the before-tax payoff is known.
rows = transpose(flipgrid)
near = (v, xs) => xs.reduce((a, b) => Math.abs(b - v) < Math.abs(a - v) ? b : a)
lookup = function (gain, lvr, rate, io) {
  const g = near(gain, [...new Set(rows.map(d => d.gain))]);
  const l = near(lvr,  [...new Set(rows.map(d => d.lvr))]);
  const r = near(rate, [...new Set(rows.map(d => d.rate))]);
  return rows.find(d => d.gain === g && d.lvr === l && d.rate === r && d.io === io);
}
```

:::: {.columns}
::: {.column width="38%"}
::: {style="font-size:0.62em"}
```{ojs}
//| echo: false
viewof h_cg   = Inputs.range([0.1, 1],     {value: 0.4, step: 0.1,    label: "Capital gain"})
viewof h_lvr  = Inputs.range([0, 1],       {value: 1, step: 0.1,      label: "LVR"})
viewof h_rate = Inputs.range([0.04, 0.09], {value: 0.06, step: 0.005, label: "Mortgage rate"})
viewof h_tau  = Inputs.select([0.15, 0.25, 0.35, 0.45, 0.47],
                              {value: 0.45, label: "Marginal tax rate",
                               format: (x) => (100 * x).toFixed(0) + "%"})
viewof h_delta = Inputs.range([0, 0.5], {value: 0.5, step: 0.05, label: "CGT discount"})
viewof h_io   = Inputs.toggle({value: true, label: "Interest only"})
```

```{ojs}
//| echo: false
md`R sent **${rows.length}** precomputed rows. The browser picks one and applies tax. Sliders snap to the grid, which is the price of moving the model out.`
```
:::
:::

::: {.column width="60%" .flip-chart}
```{ojs}
//| echo: false
h_row = lookup(h_cg, h_lvr, h_rate, h_io)

// R sent the before-tax payoff and the taxable gain; the schedule is applied
// here. Rental total is whatever the payoff is not explained by the gain.
h_out = {
  const rentalTotal = h_row.e_pre - h_row.cg_net;
  const other = incomeAtRate(h_tau);
  const rentalYr = dollars(rentalTotal / 10);
  const taxBase = scheduleTax(other);
  const taxRent = scheduleTax(other + rentalYr);
  const taxAt = (e) => (10 * (taxRent - taxBase) +
    scheduleTax(other + rentalYr + dollars((1 - h_delta) * (e - rentalTotal))) -
    taxRent) / PRICE * 100;
  const yAt = (eAnnual) => (eAnnual * 10 - taxAt(eAnnual * 10)) / 10;
  return {eYr: h_row.e_pre / 10, aYr: (h_row.e_pre - taxAt(h_row.e_pre)) / 10,
          yAt: yAt, floorYr: flipFloor(yAt)};
}

// The same drawing function as version 2 — only the model behind it changed.
flipPlot(h_out)
```
:::
::::

::: aside
`ojs_define()` is built into Quarto: it serialises an R object into the page for
Observable to read. The economics now lives in an R function you could unit-test
and show a referee, and the browser does two lines of arithmetic. This chart
calls version 2's `flipPlot()` unchanged. The cost is the grid — five parameters
at ten steps each is 100,000 rows.
:::

```{r}
#| label: frames
#| include: false

# After-tax return as a function of before-tax return, under the schedule.
after_tax <- function(e_annual, rental_total, tau, delta) {
  other     <- income_at_rate(tau)
  rental_yr <- dollars(rental_total / 10)
  base <- schedule_tax(other)
  rent <- schedule_tax(other + rental_yr)
  e    <- e_annual * 10
  gain <- schedule_tax(other + rental_yr + dollars((1 - delta) * (e - rental_total))) - rent
  (e - (10 * (rent - base) + gain) / PRICE * 100) / 10
}

frame_plot <- function(delta, tau = 0.45) {
  spell <- flip_pre_tax(0.40, 1.00, 0.06, TRUE)
  rental_total <- spell$e_pre - spell$cg_net
  curve <- tibble(e = seq(-4, 4, length.out = 97)) |>
    mutate(a = map_dbl(e, after_tax, rental_total, tau, delta))
  spell_pt <- tibble(x = spell$e_pre / 10,
                     y = after_tax(spell$e_pre / 10, rental_total, tau, delta))
  floor <- curve |> filter(a > 0) |> slice_min(e, n = 1) |> pull(e)
  band  <- if (length(floor) && floor < 0) floor else 0

  ggplot(curve, aes(e, a)) +
    annotate("rect", xmin = band, xmax = 0, ymin = -4, ymax = 4,
             fill = pal$gold, alpha = 0.18) +
    geom_hline(yintercept = 0, colour = pal$subtle, linewidth = 0.3) +
    geom_vline(xintercept = 0, colour = pal$subtle, linewidth = 0.3) +
    geom_abline(slope = 1, intercept = 0, colour = pal$subtle, linetype = "22") +
    geom_line(colour = pal$accent, linewidth = 1) +
    geom_point(data = spell_pt, aes(x, y), colour = pal$spell, size = 3) +
    annotate("text", x = 3.2, y = 2.6, label = "No tax",
             colour = pal$subtle, size = 3.4) +
    coord_cartesian(xlim = c(-4, 4), ylim = c(-4, 4)) +
    labs(title = str_glue("CGT discount of {100 * delta}%"),
         subtitle = "After-tax against before-tax return, % a year",
         x = "Before tax", y = "After tax") +
    theme_minimal(base_size = 13)
}

frames <- tibble(delta = seq(0, 0.5, by = 0.05)) |>
  mutate(path = str_c("figs/frame-", str_pad(row_number(), 2, pad = "0"), ".png"),
         shown = if_else(delta == max(delta), "frame-on", ""))

pwalk(frames, \(delta, path, shown)
      ggsave(path, frame_plot(delta), width = 6.6, height = 5.2, dpi = 150))
```

## 4. Precomputed frames from ggplot

:::: {.columns}
::: {.column width="38%"}
The chart is an ordinary ggplot, rendered eleven times in R at eleven values of
the CGT discount and stacked in the page. The slider changes which one is
visible. Nothing is computed in the browser.

```{r}
#| label: frame-controls
#| echo: false
#| results: asis
cat(str_glue(
  '<p>CGT discount of <span id="frame-out">50%</span> ',
  '<input id="frame-slider" type="range" min="1" ',
  'max="{nrow(frames)}" step="1" value="{nrow(frames)}"></p>'
))
```
:::

::: {.column width="60%"}
```{r}
#| label: frame-stack
#| echo: false
#| results: asis
frames |>
  mutate(tag = str_glue(
    '<img src="{path}" class="{shown}" alt="After-tax against before-tax ',
    'return at a CGT discount of {100 * delta} per cent.">'
  )) |>
  pull(tag) |>
  str_flatten("\n") |>
  (\(imgs) cat('<div class="frame-stack">\n', imgs, '\n</div>\n', sep = ""))()

cat('<script>
(function () {
  var s = document.getElementById("frame-slider");
  var out = document.getElementById("frame-out");
  var imgs = document.querySelectorAll(".frame-stack img");
  s.addEventListener("input", function () {
    var i = parseInt(s.value, 10) - 1;
    imgs.forEach(function (im, j) { im.classList.toggle("frame-on", j === i); });
    out.textContent = (i * 5) + "%";
  });
})();
</script>')
```
:::
::::

::: aside
All the code is R and there is no model in the browser at all. The limit is
arithmetic: one parameter at eleven steps is eleven images and about a megabyte;
two parameters is 121 images. The full seven-slider version is unreachable.
:::

## What each one costs {.smaller}

Versions 1 to 3 are pixel-identical, so only the code differs.

| | Lines | Reads like | Reusable | Ceiling |
|---|---|---|---|---|
| Hand-written SVG | ~270 | nothing legible | no | none; any effect is available |
| Observable Plot | ~30 model, ~55 spec | a chart spec | yes, it is a function | weak on history-dependent state |
| R + `ojs_define` | ~35 R, ~10 JS | an R function | yes, same spec | grid size |
| ggplot frames | ~30 R | an R script | yes, just ggplot | one parameter |

::: {.fragment}
Not here: `shinylive`, which needs a Quarto extension, ships tens of megabytes
and re-renders the whole ggplot on every tick; and Vega-Lite, whose slider
bindings are close to ideal until you need a loop, which the amortisation
schedule does.
:::

::: aside
`quarto preview demo.qmd`. Needs Quarto 1.4+, R and `tidyverse`. Slides 3 and 4
use Observable, which is blocked over `file://` and must be served.
:::
